---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(dada2)
pass <- function(x){x}
```

```{r}
packageVersion("dada2")
```


To download the data, you should run the script `PullFastqFromSRA.sh`
To download the silva daa run PullSilvaData.sh

```{bash}
sh Scripts/PullFastqFromSRA.sh
```

```{bash}
sh Scripts/PullSilvaData.sh
```


Data from SRA have serial numbers that are different from the file names to which I am accustomed.

Pull in metadata

```{r}
SRA_Data <- read_csv("StartingData/SraRunTable.csv")
```

Make some tables that are useful for data processing
```{r}
keyA1 <- SRA_Data %>% select(JacobName = `Library Name`, SRAName = Run) %>%
  separate(JacobName, into = c(NA, "Project", "Strain", "Treatment", "Group", "Run"), sep = "_", remove = FALSE)
keyA2 <- keyA1 %>%
  expand_grid(ReadDir = c(1,2)) %>%
  mutate(JacobName2 = paste0(JacobName, "_R", ReadDir),
         SRAName2 = paste0(SRAName, "_", ReadDir),
         JacobFile = paste0(JacobName2, ".fastq.gz"),
         SRAFile = paste0(SRAName2, ".fastq.gz")
         )
keyA2
```

For the phyloseq objectdada
```{r}
keyA1 %>%
  select(Sample = JacobName, Project:Run) %>% 
  write_csv("UsefulData/sample_data.csv")
```


Add paths so I can tell various steps where files go and what to call them
```{r}
keyA3 <- keyA2 %>%
  mutate(TrimmedPath = str_c("Trimmed/", JacobName2, "_trimmed.fastq.gz")) %>%
  mutate(FilteredPath = str_c("Filtered/", JacobName2, "_filtered.fastq.gz")) %>%
  mutate(DerepPath = str_c("Derep/", JacobName2, "_derep.fastq.gz"))
```



For primer trimming, I need a file of sample names

```{r}
write_lines(keyA2$JacobName, "IntermediateData/samples.txt")
```

Rename all of the files

First clear the renamed directory, if it contains any files

```{r}
do.call(file.remove, list(list.files("Renamed", full.names = "TRUE")))
```
Then copy the files to the renamed directory with their new names
```{r}
didCopy <- file.copy(paste0("SRA_Readfiles/", keyA2$SRAFile),paste0("Renamed/", keyA2$JacobFile))
didCopy
```

This should be a vector of TRUEs. If it isn't its likely that not all of the SRA files downloaded.



# Run Cutadapt
I need to have cutadapt loaded before I run rstudio on my machine for this to work

The following is commented out since it takes a while, I may uncomment once I have things working for a full path.
```{bash}
for sample in $(cat IntermediateData/samples.txt)
do

    echo "On sample: $sample"

    cutadapt -a ^GTGYCAGCMGCCGCGGTAA...AAACTYAAAKRAATTGRCGG \
-A ^CCGYCAATTYMTTTRAGTTT...TTACCGCGGCKGCTGRCAC \
 -m 200 --discard-untrimmed \
    -o Trimmed/${sample}_R1_trimmed.fastq.gz -p Trimmed/${sample}_R2_trimmed.fastq.gz \
    Renamed/${sample}_R1.fastq.gz Renamed/${sample}_R2.fastq.gz \
    >> IntermediateData/cutadapt_primer_trimming_stats.txt 2>&1
done
```

## Plot Errors

We are going to plot the Wild Fed A from the non-crash project, which was run twice, once with other on-crash samples and once wiht the crash samples. And the Wild Fed A from the crash projet.
```{r}
toPlot <- keyA3 %>% 
  #filter(Strain == "Wild", Treatment == "Fed", Group == "A")
  filter(Strain == "Wild", Treatment == "Fed", Group %in% c("A", "B"))
toPlotFwd <- toPlot %>%
  filter(ReadDir == 1) %>%
  pull(TrimmedPath)
toPlotRev <- toPlot %>%
  filter(ReadDir == 2) %>%
  pull(TrimmedPath)
```

Forward reads
```{r, fig.width = 8, fig.height = 4}
plotQualityProfile(toPlotFwd) +theme(strip.text = element_text(size = 7))
```
These are, form left to right. Non-crash sample, first run. Non crash sample final run (with crash samples). Crash sample, final run (with crash samples).
25 July 2021 -- The different samples within the two runs have remarkably similar looking error profiles, though different read numbers. Subtle differences are apparent.

Reverse reads

```{r, fig.width = 8, fig.height = 4}
plotQualityProfile(toPlotFwd) +theme(strip.text = element_text(size = 7))
```

# Filtering files

We can filter everything in one batch, but we have to do dada2 and learn errors on two batches, one for each run.

Based on the above proviles, I'm going to allow three errors on the forward read and five on the reverse read.

First, I need to get the names of the forward and reverse read files

```{r}
files_to_filter <- keyA3 %>%
  select(Project:ReadDir, TrimmedPath, FilteredPath) %>%
  #filter(Strain == "Wild", Treatment == "Fed", Group == "A") %>% # Debugging
  pivot_longer(cols = contains("Path"), names_to = "PathType", values_to = "Path") %>%
  mutate(cname = str_c(PathType, "_R", ReadDir)) %>%
  select(-PathType, -ReadDir) %>%
  pivot_wider(names_from = cname, values_from = "Path") %>%
  pass
```

Do the filtering

```{r}
pt0 <- proc.time()
filtered <- filterAndTrim(fwd  = files_to_filter$TrimmedPath_R1,
                          filt = files_to_filter$FilteredPath_R1,
                          rev  = files_to_filter$TrimmedPath_R2,
                          filt.rev  = files_to_filter$FilteredPath_R2,
                          maxEE=c(3,5),
                          rm.phix=TRUE, minLen=175, truncLen=c(230,220))
pt1 <- proc.time()
pt1 - pt0
```

```{r}
write_csv(keyA3, "IntermediateData/keyA3.csv")
write_csv(files_to_filter , "IntermediateData/files_for_learning_errors.csv")
```

Then run LearningErrors.R

```{bash}
Rscript Scripts/TestScript.R
```


```{bash}
Rscript Scripts/LearningErrors.R
```

```{bash}
Rscript Scripts/DerepAll.R
```

```{bash}
Rscript Scripts/DadaAll.R
```

```{bash}
Rscript Scripts/MergeAndTableify.R
```

```{bash}
Rscript Scripts/GetTaxa.R
```


```{bash}
Rscript Scripts/MakePhyloseqIngredients.R
```

```{bash}
Rscript Scripts/MakePhyloseqObject.R
```





